---
execute: 
  echo: false
  warning: false
---
The Air Quality in India is changing rapidly  due to rise in population in those areas. This data was sourced from a data set which was created by Kaggle. Then, different pollutants were joined to analyze the relationship between city and pollution. The example shows us that there is more air pollution in the areas with higher population.
```{r}
#| label: setup

library(tidyverse)
library(sf)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)

```

```{r}
data <- read_csv('/Users/chittibellamkonda/Downloads/archive/city_day.csv')
```

```{r}
data <- data |> drop_na()
data <- data |> distinct()
city_coords <- data.frame(
  City = c("Delhi", "Mumbai", "Kolkata", "Chennai", "Bangalore", "Hyderabad"),
  lat = c(28.7041, 19.0760, 22.5726, 13.0827, 12.9716, 17.3850),
  lon = c(77.1025, 72.8777, 88.3639, 80.2707, 77.5946, 78.4867)
)


avg_pm25 <- data |>
  group_by(City) |>
  summarise(avg_PM2.5 = mean(PM2.5, na.rm = TRUE))
india <- ne_countries(scale = "medium", country = "India", returnclass = "sf")
plot_data <- merge(avg_pm25, city_coords, by = "City")
```

```{r}
ggplot() +
  geom_sf(data = india, fill = "gray90", color = "gray50") +  # Map outline of India
  geom_point(data = plot_data, aes(x = lon, y = lat, size = avg_PM2.5, color = avg_PM2.5), alpha = 0.7) +  # PM2.5 data points
  scale_size_continuous(range = c(1, 10), name = "Average PM2.5") +  # Size scale
  scale_color_viridis_c(name = "Average PM2.5") +  # Color scale
  labs(title = "Average PM2.5 Levels by City in India") +  # Title
  theme_void() +  # Minimal theme
  theme(legend.position = "right")
```

This script maps average PM2.5 levels across major Indian cities using Kaggle's "Air Quality Data in India (2015-2020)". It helps visualize spatial differences in particulate matter levels, aiding in environmental management insights.