---
title: 'Air Quality in Delhi, India'
author: 'Akhil B'
execute: 
  echo: false
  warning: false
---
**The Air Quality in India is changing rapidly throughout the years due to the rise in population in urban areas. This data was sourced from a dataset created by Kaggle. I used this data to create graphs to see the growth of air pollution in India. Then using the graphs, different pollutants were integrated to analyze the relationship between cities and pollution. The analysis shows that there is a pattern in which the air pollution rises.**
```{r}
#| label: setup
library(gganimate)
library(tidyverse)
library(sf)
library(ggplot2)
library(viridis)
```

```{r}
# Reads The DAta
data <- read_csv('/Users/chittibellamkonda/Downloads/archive/city_day.csv')
```


```{r}
# Cleaning the data to make the city only equal to Delhi and changinng the format of
#the date
Delhi_data <- data |> 
filter(City == 'Delhi')|>
select(Date, PM2.5, PM10, NO, NO2, NOx, NH3, CO, SO2, O3)|> 
mutate(Date = as.Date(Date, format = "%Y-%m-%d"))
```

```{r}
Delhi_long <- Delhi_data |>
  pivot_longer(cols = -Date, names_to = "Pollutant", values_to = "Value") |>
  group_by(Date, Pollutant) |>
  summarise(Value = mean(Value, na.rm = TRUE))

plot_width <- 10  # Adjust as needed
plot_height <- 6    

animated_plot <- ggplot(Delhi_long, aes(x = Date, y = Value, color = Pollutant)) +
  geom_line() +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +  # Set breaks to 1 year
  labs(title = "Air Quality Metrics Over Time in Delhi",
       x = "Date", y = "Concentration",
       color = "Pollutant") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 16, face = "bold"),
        plot.margin = margin(1, 1, 3, 1, "cm"),  # Adjust margins (top, right, bottom, left)
        plot.background = element_rect(fill = "white"),
        aspect.ratio = plot_height / plot_width) +
  coord_cartesian(clip = "off")+ 
      transition_reveal(Date)  # Animate by revealing lines over time

# Save the animation as a GIF
animate(animated_plot, nframes = 200, end_pause = 50, width = plot_width, height = plot_height, units = "in", res = 100, renderer = gifski_renderer(loop = FALSE))
anim_save("hyderabad_air_quality.gif")
```


**I noticed pollutant levels tend to rise at the beginning, middle, and end of each year. This pattern likely reflects seasonal changes influenced by weather and human activities such as heating in colder months, agricultural practices like crop burning in the middle, and increased traffic during festive seasons at the end. Understanding these fluctuations is crucial for effective year-round air quality management.**