---
title: "Plot 2"
format: html
execute: 
  warning: false
  echo: false
  
---
```{r}
#| label: setup
library(gganimate)
library(tidyverse)
library(sf)
library(ggplot2)
library(viridis)
```

```{r}
# Reads The DAta
data <- read_csv('/Users/chittibellamkonda/Downloads/archive/city_day.csv')
```


```{r}
# Cleaning the data to make the city only equal to Delhi and changinng the format of
#the date
hyd_data <- data |> 
filter(City == 'Hyderabad')|>
select(Date, PM2.5, PM10, NO, NO2, NOx, NH3, CO, SO2, O3)|> 
mutate(Date = as.Date(Date, format = "%Y-%m-%d"))
```

```{r}
hyd_long <- hyd_data |>
  pivot_longer(cols = -Date, names_to = "Pollutant", values_to = "Value") |>
  group_by(Date, Pollutant) |>
  summarise(Value = mean(Value, na.rm = TRUE))

plot_width <- 10  # Adjust as needed
plot_height <- 6    

animated_plot <- ggplot(hyd_long, aes(x = Date, y = Value, color = Pollutant)) +
  geom_line() +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +  # Set breaks to 1 year
  labs(title = "Air Quality Metrics Over Time in Hyderabad",
       x = "Date", y = "Concentration",
       color = "Pollutant") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 16, face = "bold"),
        plot.margin = margin(1, 1, 3, 1, "cm"),  # Adjust margins (top, right, bottom, left)
        plot.background = element_rect(fill = "white"),
        aspect.ratio = plot_height / plot_width) +
  coord_cartesian(clip = "off")+ 
    transition_reveal(Date)  # Animate by revealing lines over time

# Save the animation as a GIF
animate(animated_plot, nframes = 200, end_pause = 50, width = plot_width, height = plot_height, units = "in", res = 100, renderer = gifski_renderer(loop = FALSE))
anim_save("hyderabad_air_quality.gif")

```
**In Hyderabad, air pollution peaks at the start and end of the year due to seasonal changes and local emissions, with PM2.5 and PM10 spiking moderately in winter. Nitrogen oxides (NO and NO2) show slight increases during colder months, while CO and SO2 are generally stable. In contrast, Delhi experiences sharp pollution spikes, especially in winter, driven by crop residue burning and vehicular emissions. PM2.5 and PM10 levels rise significantly, and NO, NO2, CO, and SO2 all increase notably during the crop burning season, highlighting more severe pollution compared to Hyderabad.**




